---
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \pagenumbering{gobble}
output:
  pdf_document:
    latex_engine: xelatex
geometry: margin=1.5cm
mainfont: Arial
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(xtable)
library(kableExtra)
```

\def\fillandplacepagenumber{%
 \par\pagestyle{empty}%
 \vbox to 0pt{\vss}\vfill
 \vbox to 0pt{\baselineskip0pt
   \hbox to\linewidth{\hss}%
   \baselineskip\footskip
   \hbox to\linewidth{%
     \hfil\thepage\hfil}\vss}}

**Supplemental Table 1:** Catalog of TotalSeqA antibodies used for study.
```{r, echo=FALSE, message=FALSE}
supTsaCatalog <- readRDS("../rds/tsa_catalog.rds")
supTsaCatalog <- supTsaCatalog %>%
  select(DNA_ID, cleanName, Clone, Barcode) %>%
  rename(Description = cleanName)

kbl(supTsaCatalog, booktabs = T, align = "c", longtable = TRUE) %>%
  kable_styling(latex_options = c("repeat_header"))
```

\pagebreak

```{r, echo=FALSE, message=FALSE}
prepareDEMarkerTable <- function(fn, as_kbl = TRUE, renameClusterCol = FALSE) {
  df <- read.csv(fn, sep = "\t")
  
  if (renameClusterCol) {
    df <- df %>%
      rename(cluster = Status)
  }
  
  df <- df %>%
    select(cluster, cleanName, avg_log2FC, p_val_adj, piScore) %>%
    mutate(p_val_adj = format(p_val_adj, scientific = TRUE, digits = 2)) %>%
    mutate(across(where(is.numeric), round, 2)) %>%
    group_by(cluster) %>%
    arrange(cluster, desc(abs(piScore))) %>%
    rename(
      Cluster = cluster,
      Marker = cleanName,
      `log2(fold change)` = avg_log2FC,
      `Adjusted p value` = p_val_adj,
      `π-score` = piScore)
  
  if (as_kbl) {
    return(kbl(df, booktabs = T, align = "c", longtable = TRUE, linesep = "") %>%
    kable_styling(latex_options = c("repeat_header")))
  } else {
    return(df)
  }
}

prepareDEMarkerTableCompiled <- function(fns, descriptors, idCol = "Subset") {
  dfs <- lapply(fns, function(f) {
    df <- prepareDEMarkerTable(f, as_kbl = FALSE, renameClusterCol = TRUE)
    return(df)
  })
  
  names(dfs) <- descriptors
  final_df <- bind_rows(dfs, .id = idCol)
  
  return(kbl(final_df, booktabs = T, align = "c", longtable = TRUE, linesep = "") %>%
    kable_styling(latex_options = c("repeat_header")))
}
```

\pagebreak

**Supplemental Table 2:** Significant surface markers enriched in each cluster (over all other cells) for *in vitro* model.
```{r, echo=FALSE, message=FALSE}
prepareDEMarkerTable("../outs/tsv/inVitro_initClusterMarkers.tsv")
```

\pagebreak

**Supplemental Table 3:** Significant surface markers enriched in HIV+ versus HIV- cells for *in vitro* model.
```{r, echo=FALSE, message=FALSE}
prepareDEMarkerTableCompiled(
  fns = c(
    "../outs/tsv/invitro_markers.tsv",
    "../outs/tsv/invitro_activeLate_markers.tsv",
    "../outs/tsv/invitro_early_markers.tsv"),
  descriptors = c(
    "CD4+",
    "activated/late diff CD4+",
    "early diff CD4+"
  )
)
```

\pagebreak
\blandscape
**Supplemental Table 4:** Differential peak list for activated CD4+ cells between HIV+ and HIV- for *in vitro* model.
```{r, echo=FALSE, message=FALSE}
ivPeakListActivated <- read.csv("../outs/tsv/invitro_markerTest_activatedPeaks_topPeaks.tsv", sep = "\t")

ivPeakListActivated <- ivPeakListActivated %>%
  mutate(Subset = ifelse(Log2FC > 0, "HIV+", "HIV-")) %>%
  select(Subset, seqnames, start, end, Log2FC, FDR, piScore, inGene, nearestTSS, nearestTSSDist) %>%
  mutate(FDR = format(FDR, scientific = TRUE, digits = 2)) %>%
  mutate(across(where(is.numeric), round, 2)) %>%
  mutate(inGene = gsub(",", ", ", inGene)) %>%
  mutate(inGene = ifelse(inGene == "FALSE", "-", inGene)) %>%
  mutate(inGene = linebreak(stringr::str_wrap(inGene, width = 15))) %>%
  rename(
    `Chr` = seqnames,
    `Start` = start,
    `End` = end,
    `π-score` = piScore,
    `In Gene` = inGene,
    `Nearest TSS Dist` = nearestTSSDist,
    `Nearest TSS` = nearestTSS
  ) %>%
  group_by(Subset) %>%
  arrange(desc(abs(`π-score`)), .by_group = TRUE)

kbl(ivPeakListActivated, booktabs = T, align = "c", longtable = TRUE, linesep = "", escape = F) %>%
    kable_styling(latex_options = c("repeat_header"))
```
\elandscape

\pagebreak

**Supplemental Table 5:** Significant surface markers enriched in each cluster (over all other cells) during chronic infection.
```{r, echo=FALSE, message=FALSE}
prepareDEMarkerTable("../outs/tsv/chronic_initClusterMarkers.tsv")
```

\pagebreak

**Supplemental Table 6:** Significant surface markers enriched in HIV+ versus HIV- cells during chronic infection.
```{r, echo=FALSE, message=FALSE}
prepareDEMarkerTableCompiled(
  fns = c(
    "../outs/tsv/chronic_markers.tsv",
    "../outs/tsv/chronic_tfh_markers.tsv"),
  descriptors = c(
    "CD4+",
    "Tfh CD4+"
  )
)
```

\pagebreak

**Supplemental Table 7:** Significant surface markers enriched in each cluster (over all other cells) during treated infection.
```{r, echo=FALSE, message=FALSE}
prepareDEMarkerTable("../outs/tsv/art_initClusterMarkers.tsv")
```

\pagebreak

**Supplemental Table 8:** Significant surface markers enriched in HIV+ versus HIV- cells during treated infection.
```{r, echo=FALSE, message=FALSE}
prepareDEMarkerTableCompiled(
  fns = c(
    "../outs/tsv/art_markers.tsv"),
  descriptors = c(
    "CD4+"
  )
)
```

\pagebreak
**Supplemental Table 9:** Autologous virus sequences used in study.
```{r, echo=FALSE, message=FALSE}
viralSeqs <- read.csv("../viralFastas/infoForSupplemental.csv", sep = ",", check.names = FALSE)

kbl(viralSeqs, booktabs = T, align = "c", longtable = TRUE) %>%
  kable_styling(latex_options = c("repeat_header"))
```







